---
title: "rSyntaxFileHomeWork2"
author: "Ronaldlee Ejalu"
date: "4/19/2021"
output:
  word_document: default
  html_document: default
---

# load the required packages
```{r}
library(ggplot2)
library(maps)
library(dplyr)
```

# set the working directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork2")
```

# load the data set
```{r}
foodservices <- read.delim(file = "../HomeWork2/datasets/FoodSrvcByCounty.txt", 
                           header = TRUE, 
                           dec = ","
)
```

# load the states data set
```{r}
statesabbr <- read.csv(file = "../HomeWork2/datasets/csvData.csv", 
                       header = TRUE, 
                       sep = ",", 
                       dec = ".")
```


# rename the columns in statesabbr data set
```{r}
names(statesabbr)[1] <- "state"
names(statesabbr)[2] <- "abbrev"
names(statesabbr)[3] <- "code"
```

# remove missing values
```{r}
foodservices <- na.omit(foodservices)
```

# column names of food services
```{r}
str(foodservices)
```



# Transform the data into the lower case
```{r}
statesabbr1 <- statesabbr %>%
  mutate(state = tolower(state), abbrev = tolower(abbrev), code = tolower(code))
```

# Returns string without leading or trailing white space
```{r}
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```


# transform the county and state code into lower case
```{r}
foodservices <- foodservices %>% 
  filter(State != "") %>%
  mutate(County = tolower(County), 
         State = tolower(State)
  ) %>%
  mutate(County = trim(County)
         , State = trim(State)
  )
```

# rename some column in the food services data set
```{r}
names(foodservices)[1] <- "county"
names(foodservices)[2] <- "code"
```


# merge the statesabbr1 data set with food services data set 
# to lookup the full name of the state
```{r}
foodservicesmergedStates <- merge(foodservices, 
                                  statesabbr1, 
                                  by.x = "code", 
                                  by.y = "code", 
                                  na.rm = TRUE
                                  )
```

# Define arguments for reshaping data
# using the build_longer_spec from the tidyr package
# use pivot_longer_spec with spec data set to transform the data
# by unpivoting
```{r}
library(tidyverse)
foodservicesSpec <- foodservicesmergedStates %>%
  tidyr::build_longer_spec(cols = starts_with("FoodServices.")
                      , names_to = "foodservicesyears"
                      , values_to = "foodservices"
                      , names_prefix = "FoodServices"
  )
#foodservicesSpec

usfoodservices <- pivot_longer_spec(foodservicesmergedStates, foodservicesSpec)
```

# transform foodservicesyears and foodservices
```{r}
usfoodservices <- usfoodservices %>% 
  mutate(foodservicesyears = trim(str_replace(foodservicesyears,"."," ")
                                  ), foodservices = as.numeric(foodservices)
         )
```

# aggregate foodservices by states in the usfoodservices data set
```{r}
foodservices.agg <- usfoodservices %>% # read foodservices data set.
#  mutate(state = as.factor(state)) %>%
  dplyr::group_by(state) %>% # group the states together
  summarise(avg.foodservices = mean(foodservices), .groups = 'drop') # find the mean 
```

# Convert the states values back to capital letters
```{r}
foodservices.agg <- foodservices.agg %>% 
  mutate(state = toupper(state)) #%>%
  #mutate(state = ifelse(state == '', 'DISTRICT OF COLUMBIA',))
```



##################################################################################
#Understaing shape files
################################################################################


# load these pacakges
```{r}
library(leaflet)
library(magrittr)
library(rgdal)
library(geojsonio)
library(htmltools)
library(htmlwidgets)
library(stringi)
library(RColorBrewer)
```

# read in the states files
```{r}
states_shp <- readOGR("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork2/States_shapefile-shp/States_shapefile.shp")
```


# determining if there is a match between join column names
```{r}
is.element(states_shp$State_Name, foodservices.agg$state)
```

# merge the shapefile with the data set
```{r}
states_merged_foodservices <- merge(states_shp, foodservices.agg, by.x = "State_Name", by.y = "state")
```

# generate a tool tip
```{r}
popup_states <- paste0("food services: ", as.character(states_merged_foodservices$avg.foodservices))
```

# create bins for the different totalfoodservices values
```{r}
bins <- c(0, 100, 200, 300, 400, 500, 600, Inf)
palbin <- colorBin("Oranges", domain = states_merged_foodservices$avg.foodservices, bins = bins) #YlGnBu
```


# generate labels
```{r}
stateLabels <- sprintf('<b>%s</b><br/>%g food services',
                       states_merged_foodservices$State_Name, states_merged_foodservices$avg.foodservices) %>%
  lapply(function(x) HTML(x))

states_merged_foodservices <- cbind(states_merged_foodservices, matrix(stateLabels, ncol = 1, dimnames = list(c(), c('stateLabels'))))
```

# This line of code NAs but it isn't working as expected
# needs to be investigated.

# ```{r}
# # FUNCTION TO REMOVE NA's IN sp DataFrame OBJECT
# #   x           sp spatial DataFrame object
# #   margin      Remove rows (1) or columns (2) 
# sp.na.omit <- function(x, margin=1) {
#   if (!inherits(x, "SpatialPointsDataFrame") & !inherits(x, "SpatialPolygonsDataFrame")) 
#     stop("MUST BE sp SpatialPointsDataFrame OR SpatialPolygonsDataFrame CLASS OBJECT") 
#   na.index <- unique(as.data.frame(which(is.na(x@data),arr.ind=TRUE))[,margin])
#     if(margin == 1) {  
#       cat("DELETING ROWS: ", na.index, "\n") 
#         return( x[-na.index,]  ) 
#     }
#     if(margin == 2) {  
#       cat("DELETING COLUMNS: ", na.index, "\n") 
#         return( x[,-na.index]  ) 
#     }
#  }
# ```
# 
# # delete NA's in Crash.map
# ```{r}
# states_merged_foodservices.noNAs <- sp.na.omit(states_merged_foodservices)
# ```
# 
# # Show changes in the dim
# ```{r}
# dim(states_merged_foodservices)
# dim(states_merged_foodservices.noNAs)
# ```


# graph the map
```{r}

m <- leaflet(states_merged_foodservices) %>% 
  
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(lng = -96.25, lat = 39.50, zoom = 4) %>%
  addPolygons(color = "grey", 
              weight = 0.2, 
              smoothFactor = 0.2, 
              fillOpacity = 0.7, 
              fillColor =  ~palbin(states_merged_foodservices$avg.foodservices), 
              popup = ~popup_states, label = ~stateLabels, 
              labelOptions = labelOptions(
                style = list(color = 'gray30'),
                textsize = '10px'),
              highlightOptions = highlightOptions(
                weight = 3,
                color = 'dodgerblue' # change this color
                )
              ) %>%
  addLegend(pal = palbin, values = states_merged_foodservices$avg.foodservices, position = "bottomright", title = "Average Food<br/>Services<br/> per<br/>U.S States")
m
```

#Strategies to removing missing values
```{r}
states_merged_foodservices@data %>% slice(9)
```


##########################################################################################################################################################
# food services by county
#########################################################################################################################################################

# load the us counties shape file
```{r}
us_counties.shp <- readOGR("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork2/cb_2019_us_county_500k/cb_2019_us_county_500k.shp")
```

# Convert the first letter of every value in the county variable to Uppercase
```{r}
usfoodservices <- usfoodservices %>%
  mutate(county = str_to_title(county))
```

# Remove Alaska(2), Hawaii(15), Puerto Rico (72), Guam (66), Virgin Islands (78), American Samoa (60)
#  Mariana Islands (69), Micronesia (64), Marshall Islands (68), Palau (70), Minor Islands (74)
```{r}
us_counties.shp <- us_counties.shp[!us_counties.shp$STATEFP %in% c("02", "15", 
                                                                   "72", "66", 
                                                                   "78", "60", 
                                                                   "69","64", 
                                                                   "68", "70", 
                                                                   "74"),]
```

## Make sure other outling islands are removed.
```{r}
us_counties.shp <- us_counties.shp[!us_counties.shp$STATEFP %in% c("81", "84", 
                                                                   "86", "87", 
                                                                   "89", "71", 
                                                                   "76", "95", "79"),]
```


#determining if there is a match between join column names of different data sets
```{r}
is.element(us_counties.shp$NAME, usfoodservices$county)
```

# aggregate foodservices by us_counties
```{r}
usfoodservicesbycounties.agg <- usfoodservices %>% # read foodservices data set.
  dplyr::group_by(county) %>% # group the states together
  summarise(avg.foodservices = mean(foodservices), 
            .groups = 'drop'
            ) # find the mean or the average
```

# Merge the us_counties.shp with the usfoodservicesbycounties data set
```{r}
us_counties.shp_merged_datafile <- merge(us_counties.shp, usfoodservicesbycounties.agg, by.x = "NAME", by.y = "county", rm.na = TRUE)
```

# Map colors to continous values
```{r}
#paletteNum <- colorNumeric("YlOrRd", domain = us_counties.shp_merged_datafile$totalfoodservices)
#paletteNum <- colorNumeric("Oranges", NULL, n= 20) #domain = us_counties.shp_merged_datafile$totalfoodservices)
bins <- c(0, 100, 200, 300, 400, 500, 600, Inf)
paletteNum <- colorBin("Oranges", domain = states_merged_foodservices$avg.foodservices, bins = bins) #YlGnBu
```



# Format popup data for leaflet map.
```{r}
popup_dat <- paste0("<strong>County: </strong>", 
                    us_counties.shp_merged_datafile$NAME, 
                    "<br><strong>Value: </strong>", 
                    us_counties.shp_merged_datafile$avg.foodservices)
```

# create county labels
```{r}
countyLabels <- sprintf('County: <b>%s</b><br/>%g food services',
                       us_counties.shp_merged_datafile$NAME, us_counties.shp_merged_datafile$avg.foodservices) %>%
  lapply(function(x) HTML(x))

states_merged_foodservices <- cbind(us_counties.shp_merged_datafile, matrix(countyLabels, ncol = 1, dimnames = list(c(), c('countyLabels'))))
```


# This line of code NAs but it isn't working as expected
# needs to be investigated.
```{r}
 # FUNCTION TO REMOVE NA's IN sp DataFrame OBJECT
 #   x           sp spatial DataFrame object
 #   margin      Remove rows (1) or columns (2)
 sp.na.omit <- function(x, margin=1) {
   if (!inherits(x, "SpatialPointsDataFrame") & !inherits(x, "SpatialPolygonsDataFrame"))
     stop("MUST BE sp SpatialPointsDataFrame OR SpatialPolygonsDataFrame CLASS OBJECT")
   na.index <- unique(as.data.frame(which(is.na(x@data),arr.ind=TRUE))[,margin])
     if(margin == 1) {
      cat("DELETING ROWS: ", na.index, "\n")
         return( x[-na.index,]  )
     }
     if(margin == 2) {
       cat("DELETING COLUMNS: ", na.index, "\n")
         return( x[,-na.index]  )
     }
  }
```
 # delete NA's in Crash.map

```{r}
 states_merged_foodservices.noNAs <- sp.na.omit(states_merged_foodservices)
```




 # Show changes in the dim
 ```{r}
 dim(states_merged_foodservices)
 dim(states_merged_foodservices.noNAs)
 ```



# Graph the us counties map
```{r}
m <- leaflet() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(lng = -96.25, lat = 39.50, zoom = 4) %>%
  addPolygons(data =  states_merged_foodservices.noNAs, 
              weight = 0.2, 
              smoothFactor = 0.5, 
              fillOpacity = 0.7, 
              fillColor =  ~paletteNum( states_merged_foodservices.noNAs$avg.foodservices), 
              popup = ~popup_dat, label = ~countyLabels, 
              labelOptions = labelOptions(
                style = list(color = 'gray30'),
                textsize = '10px'),
              highlightOptions = highlightOptions(
                weight = 3,
                color = '#2c7fb8' # change this color dodgerblue
                )
              )%>%
  addLegend(pal = paletteNum, values =  states_merged_foodservices.noNAs$avg.foodservices, position = "bottomright", title = "Average Food<br/>Services<br/> per<br/>U.S Counties")
m
```


######################################################################################################################
#Question 2 Part a 
######################################################################################################################



# set to the working directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork2")
```


# basic example - default layers
```{r}
m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lat=41.878930, lng= -87.625940, popup="DePaul CDM")
m  # Print the map
```

# read shape file from :
#https://gis.chicagopolice.org/datasets/police-beat-boundary-view-1?geometry=-93.859%2C41.124%2C-82.121%2C42.556
# do not use the KML files - they do not seem to be correctly formatted and will not keep 
# the attributes we need
# this reads the boundaries of each beat as a list of points making up a polygon
```{r}
myShape4 <- readOGR("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork2/Police_Beat_Boundary_View-shp")
myShape4@data$BEAT_NUMBE # show list of beats (still factor here)
myShape4@polygons[[1]] # show first shape in file
```

# note that 'BEAT_NUMBER' name gets truncated
# change the factor to int to match the crashes file
# could also tell readOGR to not read strings as factors
```{r}
myShape4@data$BEAT_OF_OCCURRENCE = as.integer(as.character(myShape4@data$BEAT_NUMBE))
```

# read the crashes file - which includes 'beat of occurrence' which is the 
# key used to join to the BEAT in the shape file

```{r}
crashes <- read.csv("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork2/datasets/chicago_crashes.csv")
```

#aggregate the total fatalities by beat in the crashes data set.
```{r}

crashes.agg <- crashes %>% # read crashes
  mutate(INJURIES_TOTAL = ifelse( is.na(INJURIES_TOTAL), 0, INJURIES_TOTAL)) %>% #change NA to 0
    dplyr::group_by(BEAT_OF_OCCURRENCE) %>% #group the beats together
    summarise(total = sum(INJURIES_TOTAL), .groups = 'drop') # sum up to a total
```

#this merge will add these attributes to your spatial data frame - it should be aggregated to
#the same level - i.e. 1 row per shape or less (see above)

```{r}
crash.map <- merge(myShape4, crashes.agg, 
                   by.x  = "BEAT_OF_OCCURRENCE", by.y="BEAT_OF_OCCURRENCE")
```

#see the attributes in the spacial data set
```{r}
crash.map@data$total
```


#create full leaflet map with color by total fatalities
#create palette - this gives you back a function that will later take in an integer
# and return an RGB color from a precalculated range given by the domain passed in
# I do a unique() just to reduce the # going in but it will also work with the full list
```{r}
pal <- colorNumeric(
  palette = "#2b8cbe",
  domain = unique(crash.map@data$total)) 
```

# create bins for the different total injuries
```{r}
bins <- c(0, 5, 10, 15,  Inf)
palbin <- colorBin("PuBu", domain = crash.map$total, bins = bins)
```

#generate Beat Occurence label
```{r}
beatLabels <- sprintf('<b>Beat Occurrence: %s</b><br/>Injuries total: %g',
                       crash.map$BEAT_OF_OCCURRENCE, crash.map$total) %>%
  lapply(function(x) HTML(x))

states_merged_foodservices <- cbind(crash.map, matrix(beatLabels, ncol = 1, dimnames = list(c(), c('beatLabels'))))
```

# generate a tool tip
```{r}
popup_chicago <- paste0("Injuries total: ", as.character(crash.map$total))
```



# create map with continous colors for total fatalities for whole data set
# this includes the default leaflet layer
```{r}
m <- leaflet(crash.map) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(weight = 1, smoothFactor = 0.5,
              #color = ~pal(total),    
              opacity = 1.0, 
              fillOpacity = 0.7, 
              fillColor = ~palbin(crash.map$total), 
              popup = ~popup_chicago, 
              label = beatLabels, 
              labelOptions = labelOptions(style = list(color = 'gray30')
                                            , textsize = '10px'
                                            ), 
              highlightOptions = highlightOptions(weight = 3, 
                                                    color = '#feb24c'
                                                    )
              ) %>%
  addLegend(pal = pal, values = crash.map$total, position = "bottomright", title = "Total Injuries per<br/>Beat of Occurrence")
m  # Print the map
```


##############################################################################################################
# 2b
##################################################################################################################

# load library
```{r}
library(tidyverse)
```


# create 4 bins of width 6

```{r}
crash.withbins <- crashes %>%
  mutate(bin = cut(CRASH_HOUR, seq(min(CRASH_HOUR), max(CRASH_HOUR) + 4, 6 ), right = FALSE))
```

# extract beat of occurrences that occurred between 12am and 6am 

```{r}
crashes.agg.12amto6am <- crash.withbins %>% 
  filter(bin == '[0,6)') %>%
 mutate(INJURIES_TOTAL = ifelse( is.na(INJURIES_TOTAL), 0, INJURIES_TOTAL)) %>% #change NA to 0
    dplyr::group_by(BEAT_OF_OCCURRENCE) %>% #group the beats together
    summarise(total = sum(INJURIES_TOTAL), .groups = 'drop') # sum up to a total
```

# this merge will add these attributes to your spatial data frame - it should be aggregated to
# the same level - i.e. 1 row per shape or less (see above)

```{r}
crash.map <- merge(myShape4, crashes.agg.12amto6am, 
                   by.x  = "BEAT_OF_OCCURRENCE", by.y="BEAT_OF_OCCURRENCE")
```


#create full leaflet map with color by total fatalities
#create palette - this gives you back a function that will later take in an integer
# and return an RGB color from a precalculated range given by the domain passed in
# I do a unique() just to reduce the # going in but it will also work with the full list
```{r}
pal <- colorNumeric(
  palette = "#2b8cbe",
  domain = unique(crash.map@data$total)) 
```

# create bins for the different total injuries
```{r}
bins <- c(0, 5, 10, 15,  Inf)
palbin <- colorBin("PuBu", domain = crash.map$total, bins = bins)
```

#generate Beat Occurence label
```{r}
beatLabels <- sprintf('<b>Beat Occurrence: %s</b><br/>Injuries total: %g',
                       crash.map$BEAT_OF_OCCURRENCE, crash.map$total) %>%
  lapply(function(x) HTML(x))

states_merged_foodservices <- cbind(crash.map, matrix(beatLabels, ncol = 1, dimnames = list(c(), c('beatLabels'))))
```

# generate a tool tip
```{r}
popup_chicago <- paste0("Injuries total: ", as.character(crash.map$total))
```

# Display NA rows in the crash.map 
```{r}
crash.map@data[!complete.cases(crash.map@data),]
```

```{r}
# FUNCTION TO REMOVE NA's IN sp DataFrame OBJECT
#   x           sp spatial DataFrame object
#   margin      Remove rows (1) or columns (2) 
sp.na.omit <- function(x, margin=1) {
  if (!inherits(x, "SpatialPointsDataFrame") & !inherits(x, "SpatialPolygonsDataFrame")) 
    stop("MUST BE sp SpatialPointsDataFrame OR SpatialPolygonsDataFrame CLASS OBJECT") 
  na.index <- unique(as.data.frame(which(is.na(x@data),arr.ind=TRUE))[,margin])
    if(margin == 1) {  
      cat("DELETING ROWS: ", na.index, "\n") 
        return( x[-na.index,]  ) 
    }
    if(margin == 2) {  
      cat("DELETING COLUMNS: ", na.index, "\n") 
        return( x[,-na.index]  ) 
    }
 }
```

# delete NA's in Crash.map
```{r}
crash.mapwithoutNAs <- sp.na.omit(crash.map)
```

# Show changes in the dim
```{r}
dim(crash.map)
dim(crash.mapwithoutNAs)
```


# create map with continous colors for total fatalities for whole data set
# this includes the default leaflet layer
```{r}
m <- leaflet(crash.mapwithoutNAs) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  #setView(lng = -87.6298, lat = 47.8781, zoom = 4) %>%
  addPolygons(weight = 1, smoothFactor = 0.5,
              opacity = 1.0, 
              fillOpacity = 0.7, 
              fillColor = ~palbin(crash.mapwithoutNAs$total), 
              popup = ~popup_chicago, 
              label = beatLabels, 
              labelOptions = labelOptions(style = list(color = 'gray30')
                                            , textsize = '10px'
                                            ), 
              highlightOptions = highlightOptions(weight = 3, 
                                                    color = '#feb24c'
                                                    )
              ) %>%
  addLegend(pal = palbin, values = crash.mapwithoutNAs$total, position = "bottomright", title = " 12am to 6am<br/>Total Injuries per<br/> Beat of Occurrence", opacity = 1)
m  # Print the map
```

########################################################
# 6 am to 12 pm
##########################################################


# extract beat of occurrences that occurred between 6am and 12pm 
# and sum up Injuries total and group the beats together

```{r}
crashes.agg.6amto12pm <- crash.withbins %>% 
  filter(bin == '[6,12)') %>%
 mutate(INJURIES_TOTAL = ifelse( is.na(INJURIES_TOTAL), 0, INJURIES_TOTAL)) %>% #change NA to 0
    dplyr::group_by(BEAT_OF_OCCURRENCE) %>% #group the beats together
    summarise(total = sum(INJURIES_TOTAL), .groups = 'drop') # sum up to a total
```

# this merge will add these attributes to your spatial data frame - it should be aggregated to
# the same level - i.e. 1 row per shape or less (see above)

```{r}
crash.map <- merge(myShape4, crashes.agg.6amto12pm, 
                   by.x  = "BEAT_OF_OCCURRENCE", by.y="BEAT_OF_OCCURRENCE")
```


#create full leaflet map with color by total fatalities
#create palette - this gives you back a function that will later take in an integer
# and return an RGB color from a precalculated range given by the domain passed in
# I do a unique() just to reduce the # going in but it will also work with the full list
```{r}
pal <- colorNumeric(
  palette = "#2b8cbe",
  domain = unique(crash.map@data$total)) 
```

# create bins for the different total injuries
```{r}
bins <- c(0, 5, 10, 15,  Inf) #  RdYlGn, 2PuBuGn
palbin <- colorBin("PuBu", domain = crash.map$total, bins = bins)
```

#generate Beat Occurence label
```{r}
beatLabels <- sprintf('<b>Beat Occurrence: %s</b><br/>Injuries total: %g',
                       crash.map$BEAT_OF_OCCURRENCE, crash.map$total) %>%
  lapply(function(x) HTML(x))

states_merged_foodservices <- cbind(crash.map, matrix(beatLabels, ncol = 1, dimnames = list(c(), c('beatLabels'))))
```

# generate a tool tip
```{r}
popup_chicago <- paste0("Injuries total: ", as.character(crash.map$total))
```


# Display NA rows in the crash.map 
```{r}
crash.map@data[!complete.cases(crash.map@data),]
```

```{r}
# FUNCTION TO REMOVE NA's IN sp DataFrame OBJECT
#   x           sp spatial DataFrame object
#   margin      Remove rows (1) or columns (2) 
sp.na.omit <- function(x, margin=1) {
  if (!inherits(x, "SpatialPointsDataFrame") & !inherits(x, "SpatialPolygonsDataFrame")) 
    stop("MUST BE sp SpatialPointsDataFrame OR SpatialPolygonsDataFrame CLASS OBJECT") 
  na.index <- unique(as.data.frame(which(is.na(x@data),arr.ind=TRUE))[,margin])
    if(margin == 1) {  
      cat("DELETING ROWS: ", na.index, "\n") 
        return( x[-na.index,]  ) 
    }
    if(margin == 2) {  
      cat("DELETING COLUMNS: ", na.index, "\n") 
        return( x[,-na.index]  ) 
    }
 }
```

# delete NA's in Crash.map
```{r}
crash.mapwithoutNAs <- sp.na.omit(crash.map)
```

# Show changes in the dim
```{r}
dim(crash.map)
dim(crash.mapwithoutNAs)
```


# create map with continous colors for total fatalities for whole data set
# this includes the default leaflet layer
```{r}
m <- leaflet(crash.mapwithoutNAs) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(weight = 1, smoothFactor = 0.5,
              #color = ~pal(total),    
              opacity = 1.0, 
              fillOpacity = 0.7, 
              fillColor = ~palbin(crash.mapwithoutNAs$total), 
              popup = ~popup_chicago, 
              label = beatLabels, 
              labelOptions = labelOptions(style = list(color = 'gray30')
                                            , textsize = '10px'
                                            ), 
              highlightOptions = highlightOptions(weight = 3, 
                                                    color = '#feb24c'
                                                    )
              ) %>%
  addLegend(pal = palbin, values = crash.mapwithoutNAs$total, position = "bottomright", title = "6am to 12pm<br/>Total Injuries per<br/> Beat of Occurrence", opacity = 1)
m  # Print the map
```


################################################################################################################################
#12pm to 18pm
################################################################################################################################


# extract beat of occurrences that occurred between 6am and 12pm 
# and sum up Injuries total and group the beats together

```{r}
crashes.agg.12pmto18pm <- crash.withbins %>% 
  filter(bin == '[12,18)') %>%
 mutate(INJURIES_TOTAL = ifelse( is.na(INJURIES_TOTAL), 0, INJURIES_TOTAL)) %>% #change NA to 0
    dplyr::group_by(BEAT_OF_OCCURRENCE) %>% #group the beats together
    summarise(total = sum(INJURIES_TOTAL), .groups = 'drop') # sum up to a total
```

# this merge will add these attributes to your spatial data frame - it should be aggregated to
# the same level - i.e. 1 row per shape or less (see above)

```{r}
crash.map <- merge(myShape4, crashes.agg.12pmto18pm, 
                   by.x  = "BEAT_OF_OCCURRENCE", by.y="BEAT_OF_OCCURRENCE")
```


#create full leaflet map with color by total fatalities
#create palette - this gives you back a function that will later take in an integer
# and return an RGB color from a precalculated range given by the domain passed in
# I do a unique() just to reduce the # going in but it will also work with the full list
```{r}
pal <- colorNumeric(
  palette = "#2b8cbe",
  domain = unique(crash.map@data$total)) 
```


# create bins for the different total injuries
```{r}
bins <- c(0, 5, 10, 15,  Inf) #  Option colors 10 RdYlGn, 2) PuBuGn
palbin <- colorBin("PuBu", domain = crash.map$total, bins = bins)
```

#generate Beat Occurence label
```{r}
beatLabels <- sprintf('<b>Beat Occurrence: %s</b><br/>Injuries total: %g',
                       crash.map$BEAT_OF_OCCURRENCE, crash.map$total) %>%
  lapply(function(x) HTML(x))

states_merged_foodservices <- cbind(crash.map, matrix(beatLabels, ncol = 1, dimnames = list(c(), c('beatLabels'))))
```

# generate a tool tip
```{r}
popup_chicago <- paste0("Injuries total: ", as.character(crash.map$total))
```



# create map with continous colors for total fatalities for whole data set
# this includes the default leaflet layer
```{r}
m <- leaflet(crash.map) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(weight = 1, smoothFactor = 0.5,
              #color = ~pal(total),    
              opacity = 1.0, 
              fillOpacity = 0.7, 
              fillColor = ~palbin(crash.map$total), 
              popup = ~popup_chicago, 
              label = beatLabels, 
              labelOptions = labelOptions(style = list(color = 'gray30')
                                            , textsize = '10px'
                                            ), 
              highlightOptions = highlightOptions(weight = 3, 
                                                    color = '#feb24c'
                                                    )
              ) %>%
  addLegend(pal = palbin, values = crash.map$total, position = "bottomright", title = "12pm to 6pm<br/>Total Injuries per<br/> Beat of Occurrence", opacity = 1)
m  # Print the map
```

###########################################################################################################################
#18pm to 24
############################################################################################################################


# extract beat of occurrences that occurred between 6am and 12pm 
# and sum up Injuries total and group the beats together

```{r}
crashes.agg.18pmto24 <- crash.withbins %>% 
  filter(bin == '[18,24)') %>%
 mutate(INJURIES_TOTAL = ifelse( is.na(INJURIES_TOTAL), 0, INJURIES_TOTAL)) %>% #change NA to 0
    dplyr::group_by(BEAT_OF_OCCURRENCE) %>% #group the beats together
    summarise(total = sum(INJURIES_TOTAL), .groups = 'drop') # sum up to a total
```

# this merge will add these attributes to your spatial data frame - it should be aggregated to
# the same level - i.e. 1 row per shape or less (see above)

```{r}
crash.map <- merge(myShape4, crashes.agg.18pmto24, 
                   by.x  = "BEAT_OF_OCCURRENCE", by.y="BEAT_OF_OCCURRENCE")
```



#create full leaflet map with color by total fatalities
#create palette - this gives you back a function that will later take in an integer
# and return an RGB color from a precalculated range given by the domain passed in
# I do a unique() just to reduce the # going in but it will also work with the full list
```{r}
pal <- colorNumeric(
  palette = "#2b8cbe",
  domain = unique(crash.map@data$total)) 
```

# create bins for the different total injuries
```{r}
bins <- c(0, 5, 10, 15,  Inf) #  Option colors 10 RdYlGn, 2) PuBuGn
palbin <- colorBin("PuBu", domain = crash.map$total, bins = bins)
```

#generate Beat Occurence label
```{r}
beatLabels <- sprintf('<b>Beat Occurrence: %s</b><br/>Injuries total: %g',
                       crash.map$BEAT_OF_OCCURRENCE, crash.map$total) %>%
  lapply(function(x) HTML(x))

states_merged_foodservices <- cbind(crash.map, matrix(beatLabels, ncol = 1, dimnames = list(c(), c('beatLabels'))))
```

# generate a tool tip
```{r}
popup_chicago <- paste0("Injuries total: ", as.character(crash.map$total))
```


# Display NA rows in the crash.map 
```{r}
crash.map@data[!complete.cases(crash.map@data),]
```

```{r}
# FUNCTION TO REMOVE NA's IN sp DataFrame OBJECT
#   x           sp spatial DataFrame object
#   margin      Remove rows (1) or columns (2) 
sp.na.omit <- function(x, margin=1) {
  if (!inherits(x, "SpatialPointsDataFrame") & !inherits(x, "SpatialPolygonsDataFrame")) 
    stop("MUST BE sp SpatialPointsDataFrame OR SpatialPolygonsDataFrame CLASS OBJECT") 
  na.index <- unique(as.data.frame(which(is.na(x@data),arr.ind=TRUE))[,margin])
    if(margin == 1) {  
      cat("DELETING ROWS: ", na.index, "\n") 
        return( x[-na.index,]  ) 
    }
    if(margin == 2) {  
      cat("DELETING COLUMNS: ", na.index, "\n") 
        return( x[,-na.index]  ) 
    }
 }
```

# delete NA's in Crash.map
```{r}
crash.mapwithoutNAs <- sp.na.omit(crash.map)
```

# Show changes in the dim
```{r}
dim(crash.map)
dim(crash.mapwithoutNAs)
```


# create map with continous colors for total fatalities for whole data set
# this includes the default leaflet layer
```{r}
m <- leaflet(crash.mapwithoutNAs) %>%
  #drop_na() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(weight = 1, smoothFactor = 0.5,
              #color = ~pal(total),    
              opacity = 1.0, 
              fillOpacity = 0.7, 
              fillColor = ~palbin(crash.mapwithoutNAs$total), 
              popup = ~popup_chicago, 
              label = beatLabels, 
              labelOptions = labelOptions(style = list(color = 'gray30')
                                            , textsize = '10px'
                                            ), 
              highlightOptions = highlightOptions(weight = 3, 
                                                    color = '#feb24c'
                                                    )
              ) %>%
  addLegend(pal = palbin, values = crash.mapwithoutNAs$total, position = "bottomright", title = "6pm to midnight<br/>Total Injuries per<br/> Beat of Occurrence", opacity = 1)
m  # Print the map
```


#######################################################################################################################################
# 3a Simple moving average
#######################################################################################################################################

# set the working directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork2")
```

# load the data set
```{r}
PortlandWaterLevel2003 <- read.csv(file = "../HomeWork2/datasets/PortlandWaterLevel2003.csv", 
                           header = TRUE, 
                           sep = ","
)
```

# structure of the data set
```{r}
str(PortlandWaterLevel2003)
```

# Converting the date field to date
# in the format of the 1st month of every year
```{r}
library(lubridate)
PortlandWaterLevel2003 <- PortlandWaterLevel2003 %>%
  mutate(Date = mdy(Date)
         )
```




# generate the plot
```{r}
library(tidyquant)
wlplot <-  ggplot(PortlandWaterLevel2003, aes(x= Date, y = WL)) +
  #geom_line() 
 expand_limits(y=0) # Make the y-axis go to zero.
wlplot <- wlplot + geom_ma(ma_fun = SMA, n = 100, size = 1, color = "black")
wlplot <- wlplot + theme(axis.text.x = element_text(face="bold.italic", 
                                   size=10
                                   ),
        
        axis.text.y = element_text(face="bold.italic", 
                                   size = 10
                                   ), 
        panel.grid.major.x = element_blank(), # Hide the vertical grid lines 
        panel.grid.minor.x = element_blank(),
        
        # panel.grid.major.y = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.y = element_blank(), 
        legend.position = "none", 
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
  labs(y  = "Water Level", x = "Date", title = "Moving Average of Portland Water Levels for the year 2003") +
  scale_x_date(breaks = "3 months") 
  
wlplot
```

#########################################################################################################
#Area Curve
#########################################################################################################

# Extract out the hours of the Time variable
```{r}
library(stringi)
PortlandWaterLevel2003 <- PortlandWaterLevel2003 %>%
  mutate(HrsOfWaterLevels = substring(PortlandWaterLevel2003$Time, 
                                      0, 
                                      unlist(gregexpr(pattern = ":", 
                                                      PortlandWaterLevel2003$Time
                                                      )
                                             ) - 1
                                      )  %>% as.numeric()
         )
```


# create 4 bins of 6 hrs difference

```{r}
PortlandWaterLevel2003 <- PortlandWaterLevel2003 %>%
  mutate(RangeOfHours = cut(HrsOfWaterLevels, seq(min(HrsOfWaterLevels), max(HrsOfWaterLevels) + 4, 6 ), right = FALSE))
```

# Group the Date and RangeOfHours together
# and find the average
```{r}
PortlandWaterLevel2003.agg <- PortlandWaterLevel2003 %>%
  dplyr::group_by(Date, RangeOfHours) %>%
  summarise(avgWL = mean(WL), .groups = 'drop')
```

# derive the graph
```{r}
areamap <- ggplot(PortlandWaterLevel2003.agg, aes(x = Date, y = avgWL, fill = RangeOfHours)) +
  expand_limits(y = 0) + 
  geom_area() + 
  scale_fill_viridis_d() + 
  theme_light() + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=10
                                   ),
        
        axis.text.y = element_text(face="bold.italic", 
                                   size = 10
                                   ), 
        panel.grid.major.x = element_blank(), # Hide the vertical grid lines 
        panel.grid.minor.x = element_blank()
        , plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
  labs(y  = "% Average Water Level", x = "Date", title = "Cycles of water levels that happen every day per the Hour Range") +
  scale_x_date(breaks = "3 months") 
 

areamap
```

