---
title: "Group Project"
author: "Ronaldlee Ejalu"
date: "5/11/2021"
output:
  word_document: default
  html_document: default
---

# Load the necessary libraries
```{r}
library(tidyverse)
library(dplyr)


library(ggplot2)
library(RColorBrewer)
library(scales) 
library(gifski)
library(gganimate)
```

# Set a working directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork3")
```

# load the data set
```{r}
Flightsds <- read.csv(file = "../HomeWork3/datasets/2016.csv", sep = "," , header = TRUE )
```

# View the first 10 ten records
```{r}
head(Flightsds)
```

# determine any missing values
```{r}
sum(is.na(Flightsds))
```

# structure of the data set
```{r}
str(Flightsds)
```

# lets select a sample of 100000 rows
```{r}
flight.sampleds <- Flightsds[sample(nrow(Flightsds), 100000), ]
```

# cleaning the data
```{r}
Flightsds$Unnamed..27 <- NULL 
flight.sampleds.nona <-  na.omit(Flightsds)
flight.sampleds.nona <- flight.sampleds.nona %>% 
  mutate(month = format(as.Date(FL_DATE), 
                        "%B") %>% as.factor(), 
         Carrier = as.factor(OP_CARRIER)
      ) 
```


# graph a scatter plot
```{r}
# deltaFlights <- flight.sampleds %>% filter(OP_CARRIER == "DL")

splot <- ggplot(Flightsds, aes(x=(DEP_DELAY/100), y = (ARR_DELAY/100))) +
  geom_point(alpha = 0.2, na.rm = TRUE) +
  #geom_jitter(width = 30, height=30, na.rm = TRUE) +
  theme(axis.text.x = element_text(face="bold.italic",
                                   size=12),
        axis.text.y = element_text(face="bold.italic",

                                   size=12
                                    ),
        plot.title = element_text(hjust = 0.5,
                                  size = 14,
                                  face = "bold.italic"
                                  ) # format the title
        ) +
  labs(title = " A scatter plot of Arrival Delay Vs Departure Delay for all carriers", x = "Departure Delay\n(in hundreds)", y = "Arrival Delay\n (in hundreds)") + facet_wrap(~ as.factor(OP_CARRIER))
splot
```
# There is a positive relationship between dep_delay and arr_delay
# As dep_delays increases, arrive delays tend to increase. 

#qa
```{r}
#library(lubridate)
Flightsds$month <- format(as.Date(Flightsds$FL_DATE), "%B")
```


# Number of flights of each airline in 2016
```{r}

flightsCounts <- Flightsds %>%
  mutate(dayofweek = weekdays(as.Date(FL_DATE))) %>%
  mutate(dayofweek = case_when(dayofweek == "Monday" ~ "Mon",
                               dayofweek == "Tuesday" ~ "Tue",
                               dayofweek == "Wednesday" ~ "Wed",
                               dayofweek == "Thursday" ~ "Thur",
                               dayofweek == "Friday" ~ "Fri",
                               dayofweek == "Saturday" ~ "Sat",
                               dayofweek == "Sunday" ~ "Sun"
                               )
         ) %>%
  group_by(OP_CARRIER,month, dayofweek) %>% # group the number of flights by carrier
  dplyr::summarize(numofflights = n(), .groups = "drop") %>%
  mutate(Carrier = factor(OP_CARRIER), month = as.factor(month)) %>%
  mutate(dayofweek = factor(dayofweek, 
                            levels = c("Mon", "Tue", "Wed", "Thur", "Fri", "Sat", "Sun")
                            )
         )
```

# Number of flights of each airline in 2016 with vertical axis scale modified
```{r}
library(plotly)
# flightsCounts <- Flightsds %>%
#   mutate(dayofweek = weekdays(as.Date(FL_DATE))) %>%
#   mutate(dayofweek = case_when(dayofweek == "Monday" ~ "Mon",
#                                dayofweek == "Tuesday" ~ "Tue",
#                                dayofweek == "Wednesday" ~ "Wed",
#                                dayofweek == "Thursday" ~ "Thur",
#                                dayofweek == "Friday" ~ "Fri",
#                                dayofweek == "Saturday" ~ "Sat",
#                                dayofweek == "Sunday" ~ "Sun")) %>%
#   group_by(OP_CARRIER,month, dayofweek) %>% # group the number of flights by carrier
#   summarize(numofflights = n(), .groups = "drop") %>%
#   mutate(Carrier = as.factor(OP_CARRIER), month = as.factor(month)) %>%
#   mutate(dayofweek = factor(dayofweek, levels = c("Mon", "Tue", "Wed", "Thur", "Fri", "Sat", "Sun")))

bplot <- flightsCounts %>% filter(month== "January" & Carrier == "WN") %>% ggplot( aes(x=fct_reorder(dayofweek,
                                                 (numofflights/1000000)#,
                                                 #.desc = TRUE
                                                 ),
                                   y = (numofflights/1000000), text = paste('</br>Day of Week : ', dayofweek, '</br>Number of Flights: ', (numofflights/1000000)

                ))) +
  geom_col() +
  expand_limits(y=0) +
  theme(axis.text.x = element_text(face="bold.italic",
                                   size=12),
        axis.text.y = element_text(face="bold.italic",

                                   size=12
                                    ),
        plot.title = element_text(hjust = 0.5,
                                  size = 14,
                                  face = "bold.italic"
                                  ) # format the title
        ) +
  labs(title = "January 2016 Daily Number of Flights for SouthWest Airlines", x = "Day of the Week", y = "Number of Flights\n (in Millions)") +
  #scale_y_continuous(labels = scales::comma, breaks = c(0, 0.25, 0.5, 0.75, 1.0)) +
  theme_bw()

fig <- ggplotly(bplot, tooltip = c("text"))
fig
```

# subselect all the Delay variables
```{r}
flightsdelay <- flight.sampleds %>%
  select(DEP_DELAY, ARR_DELAY, CARRIER_DELAY, WEATHER_DELAY, NAS_DELAY, SECURITY_DELAY, LATE_AIRCRAFT_DELAY)
```

# Scatter plot matrix for all the delay variables
```{r}
# library(GGally)
# ggpairs(flightsdelay, cardinality_threshold = 211,  na="naDiag", lower=list(continuous='points'),
#     axisLabels='none',
#     upper=list(continuous='blank'))
```

# Univariate Scatter plot
```{r}
#Flightsds$Unnamed..27 <- NULL 
#flight.sampleds.nona <-  na.omit(Flightsds)

bplot <-  flight.sampleds.nona %>% 
  filter(month == "January") %>% 
  ggplot(aes(x = Carrier,
                                     y = CARRIER_DELAY
                                     )
                ) +
  geom_point(alpha = 0.5) +
  #stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3 , fill = "white") +
  theme(axis.text.x = element_text(face="bold.italic",
                                   size=12),
        axis.text.y = element_text(face="bold.italic",

                                   size=12
                                    ),
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5,
                                  size = 14,
                                  face = "bold.italic"
                                  ) # format the title
        ) +
    labs(title = "Univariate Scatterplot showing Carrier Delay vs the Carriers", x = "Carriers", y = "Carrier Delay") + 
  theme_bw()


bplot
```

# create new variables
# flightStatus and statusdescr
```{r}
flight.withnewVariables <- Flightsds %>% 
  mutate(flightStatus = case_when(ARR_DELAY < 15 & ARR_DELAY > 0  ~ 0, 
                                  ARR_DELAY >= 15 & ARR_DELAY > 60 ~ 1, 
                                  ARR_DELAY >= 60 ~ 2, 
                                  DIVERTED == 1 ~ 3, 
                                  CANCELLED == 1 ~ 4
                                  )
  ) %>%
  mutate(Status = as.factor(case_when(flightStatus == 0 ~ 'On time', 
                                      flightStatus == 1 ~ 'Slighty delayed', 
                                      flightStatus == 2 ~ 'highly delayed', 
                                      flightStatus == 3 ~ 'diverted', 
                                      flightStatus == 4 ~ 'cancelled'
                                      )
                            ), 
         month = format(as.Date(FL_DATE), 
                        "%B") %>% as.factor(), 
  ) %>%
  mutate(dayofweek = weekdays(as.Date(FL_DATE))) %>% 
  mutate(dayofweek = case_when(dayofweek == "Monday" ~ "Mon", 
                               dayofweek == "Tuesday" ~ "Tue", 
                               dayofweek == "Wednesday" ~ "Wed", 
                               dayofweek == "Thursday" ~ "Thur", 
                               dayofweek == "Friday" ~ "Fri", 
                               dayofweek == "Saturday" ~ "Sat", 
                               dayofweek == "Sunday" ~ "Sun")  , 
         Carrier = as.factor(OP_CARRIER)
         ) %>%
  mutate(dayofweek = factor(dayofweek, levels = c("Mon", "Tue", "Wed", "Thur", "Fri", "Sat", "Sun")))
```

# Number of flights per status.
```{r}
cnt.agg.flighstatus <- flight.withnewVariables %>%
  group_by(Status,  month, dayofweek, Carrier) %>%
  dplyr::summarize(numflights = n(),.groups = "drop")

# any missing values
cnt.agg.flighstatus <- na.omit(cnt.agg.flighstatus)

cnt.agg.flighstatus <- cnt.agg.flighstatus %>% 
  mutate(numflights=sort(numflights, decreasing = TRUE)) # Order by num of flights in descreasing order
```

# Graph the flight status using the bar plot
```{r}
bplot <- cnt.agg.flighstatus %>% 
  filter(month == "January" & Carrier == "WN") %>% 
  ggplot(aes(x=fct_reorder(dayofweek, 
                                                 (numflights/1000)#, .desc = TRUE
                                                 ), 
                                   y = (numflights/1000), fill = Status
                                   )
                ) +
  geom_col(position = "dodge", colour = "black") + 
  expand_limits(y=0) + scale_fill_manual(values = c("#a6cee3", "#1f78b4","#b2df8a","#fbb4ae", "#d95f02")) + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
  labs(title = "January 2016 Daily Carrier Number of Flights per Status for SouthWest Airlines", x = "Day of Week", y = "Number of Flights \n in thousands") + 
  #coord_flip() + 
  scale_fill_discrete(limits = c("cancelled", "diverted", "highly delayed", "On time", "Slightly delayed")) +
  theme_bw() 

bplot
```

# Cancelled flights
# filter by flightStatus equal to 4
```{r}
CancelledFlights <- 
  flight.withnewVariables %>% 
  filter(flightStatus == 4) %>%
  mutate(month = format(as.Date(FL_DATE), "%B") %>% 
           as.factor()
         )
```

# determine the number of cancelled flights per CancellationCode
```{r}
CancelledFlights.agg <- CancelledFlights %>% 
  group_by(CANCELLATION_CODE, month, dayofweek, Carrier) %>% 
  dplyr::summarize(num.flights = n(), .groups = "drop") %>%

  mutate(Reason = as.factor(case_when(CANCELLATION_CODE == 'A' ~ 'Carrier', 
                                      CANCELLATION_CODE == 'B' ~ 'Weather', 
                                      CANCELLATION_CODE == 'C' ~ 'National Air System', 
                                      CANCELLATION_CODE == 'D' ~ 'Security'
                                      )
                            )
         ) %>% # derive the cancellation reason 
  mutate(num.flights = sort(num.flights, decreasing = TRUE)
          ) %>% # Order by num of flights
  mutate(CANCELLATION_CODE = as.factor(CANCELLATION_CODE)) #%>%
```


# derive the box for Cancellation
```{r}
#CancelledFlights.agg <- CancelledFlights.agg %>% mutate(Reason = factor(Reason, levels=c("Carrier", "Weather", "National Air System", "Security")))

bplot <- CancelledFlights.agg %>%
  filter(month == "January" & Carrier == "WN") %>%
  mutate(Reason = factor(Reason, levels=c("Carrier", "Weather", "National Air System", "Security"))) %>%
  
  ggplot(aes(x=fct_reorder(dayofweek, 
                           (num.flights/1000)
                           ), 
             y = (num.flights/1000), fill = Reason
             )
         ) +
  geom_col(position = "dodge", colour = "black") + 
  expand_limits(y=0) + 
  scale_fill_manual(values = c("#a6cee3", "#1f78b4","#b2df8a","#fbb4ae", "#d95f02"
                               )
                    ) + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
        ),  
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
        ) # format the title
  ) + 
  labs(title = "January 2016 Daily Carrier Cancelled Flights per Cancellation Reason for Southwest Airlines", x = "Day of the Week", y = "Number of Cancelled Flights \n in thousands")  + 
  scale_fill_discrete(limits=c("Carrier", "Weather", "National Air System", "Security")) + # controlling the order of the legend
  # coord_flip() + 
  theme_bw()

bplot
```

# 'ArrDelay', 'CarrierDelay', 'LateAircraftDelay', 'NASDelay', 'WeatherDelay'

```{r}
str(Flightsds)
```


# Pivot the data
flight
```{r}
Flightsds.pivot <- na.omit(Flightsds) %>% # Pivot the data
  pivot_longer(c("ARR_DELAY", "CARRIER_DELAY", "WEATHER_DELAY", "NAS_DELAY", "LATE_AIRCRAFT_DELAY"), names_to = "OtherDelays", values_to = "Values") %>% 
  mutate(OtherDelays = as.factor(OtherDelays), month = format(as.Date(FL_DATE), "%B") %>% as.factor())
```

# derive the boxplot showing the different distributions
```{r}
bplot <- Flightsds.pivot %>% 
  filter(month=="January") %>% 
  ggplot(aes(x = OtherDelays, y = Values)) + 
  geom_boxplot() + # stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3 , fill = "white") +
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
    labs(title = "Box plots comparing the distributions of Arrival, Carrier, Weather, NAS and late aircraft delays aganist their values.")  + theme_bw() + coord_flip()

# bplot <- bplot + geom_jitter()

bplot

```

# structure of flightdelay
```{r}
str(flight.sampleds)
```

# subset the data
```{r}
flights1 <- Flightsds %>% 
  select(ARR_DELAY, DEP_DELAY, FL_DATE, OP_CARRIER) %>% 
  mutate(FL_DATE = as.Date(FL_DATE), 
         OP_CARRIER = as.factor(OP_CARRIER), 
         FlightMonth = format(FL_DATE, "%m") %>% as.integer(), 
         month = format(as.Date(FL_DATE), "%B") %>% as.factor(), 
         day = format(as.Date(FL_DATE), "%d") %>% as.integer()) %>%
  group_by(ARR_DELAY, DEP_DELAY, FL_DATE, OP_CARRIER, FlightMonth, month, day) %>%
  dplyr::summarize(noOfFlights = n(), .groups = "drop")
```

# graph the animated scattter splot
```{r}
# deltaFlights <- flight.sampleds %>% filter(OP_CARRIER == "DL")

splot <- flights1 %>% 
  filter(month == "May") %>% 
  ggplot(aes(x=(DEP_DELAY/100), y = (ARR_DELAY/100), size = noOfFlights, colour = OP_CARRIER)) + 
  geom_point(show.legend = FALSE, alpha = 0.7) + 
    scale_color_viridis_d() +
  scale_size(range = c(2, 12)) + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic",  
                                   
                                   size=12
                                    ),  
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
  labs(title = " A scatter plot of Arrival Delay Vs Departure Delay for all carriers", x = "Departure Delay\n(in hundreds)", y = "Arrival Delay\n (in hundreds)") 
splot
```
# add on a layer
```{r}
splot <- splot + transition_time(day) +
  labs(title = "Flight day: {frame_time}") 
splot
```





