---
title: "RsyntaxDSC465HomeWork3"
author: "Ronaldlee Ejalu"
date: "5/6/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

# Package Installation scripts
```{r}
# library(rlang)
# sessionInfo()

#remove.packages("dplyr")
# options("install.lock"=FALSE)
# install.packages("dplyr", dependencies = TRUE, INSTALL_opts = '--no-lock')
```


# load the necessary packages
```{r}
library(dplyr)
library(tidyverse)

library(ggplot2)
library(RColorBrewer)
```


# Set a working directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork3")
```

# load the data set
```{r}
PerceptionExp <- read.csv(file = "../HomeWork3/datasets/PerceptionExperiment.csv", sep = "," , header = TRUE )
```

# number of observation of the data set
```{r}
dim(PerceptionExp)
```

# structure of the data set
```{r}
str(PerceptionExp)
```

# determine any missing values
```{r}
sum(is.na(PerceptionExp))
```

# Understanding the data
```{r}
#library(dplyr)
PerceptionExp %>% filter(Trial == "B") %>% select(Test, TestNumber, Subject)
```

# doing more exploratory analysis
```{r}
PerceptionExp %>% select(Test) %>% unique()
```

# doing more exploratory analysis
# with groupings
```{r}
PerceptionExp <- PerceptionExp %>% 
  group_by(Test, Display, TestNumber) %>%
  arrange(TestNumber, .by_group = TRUE)
```

# Create a new column that contains amount of Error
```{r}
PerceptionExp <- PerceptionExp %>% mutate(Error = Response - TrueValue, absoluteError = abs(Response - TrueValue), Test = as.factor(Test)) %>%
  mutate(estimateIndicator = case_when(Error == 0.0 ~ " Equal", 
                                       Error < 0.00 ~ "Underestimated", 
                                       Error > 0 ~ "Overestimated")
         )
```

##############################################################################################################
# 1 a grouped bar chart
#################################################################################################################

# using a grouped bar chart
```{r}
Perception.agg <- PerceptionExp %>%
  dplyr::group_by(Test, estimateIndicator) %>% # group the data by Test and EstimatedIndicator
  summarise(countoferror = n(), .groups = 'drop') %>% # find the count of error
  mutate(EstimateIndicator = as.factor(estimateIndicator))

barplot <- ggplot(Perception.agg, aes(x = reorder(factor(Test), countoferror), y = countoferror, fill = EstimateIndicator )) +
  geom_col(position = "dodge", colour = "black") +  # use the position parameter to create the grouped bar charts
  
  # This is for color blind people
  scale_fill_manual(values = c("#a6cee3", "#1f78b4","#b2df8a")) #+ scale_fill_brewer(palette = "Pastel1")


barplot <- barplot + theme(axis.text.x = element_text(face="bold.italic", 
                                   size=10, 
                                   angle=10),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=10
                                    ),  
        panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + labs(title = "A bar graph showing the comparison of equal, overestimated and underestimated responses over the different Test", 
                 x = "Test")

barplot
```


# Univariate scatter plot
```{r}
peplot <- PerceptionExp %>% ggplot(aes(x = fct_reorder(Test, absoluteError), y = absoluteError, colour = Test)) + 
  geom_point(size = 1.5) + scale_color_brewer(palette = "Dark2")

peplot <- peplot + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=10, 
                                   angle=10),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=10
                                    ),  
        panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ),  # format the title
        
        legend.background = element_rect(fill = "white", colour = "grey", size = 1), 
        legend.title = element_text(face = "bold", size = 14),
        legend.text = element_text(colour = "black"), 
        legend.key = element_rect(colour = "black", size = 0.1)
        ) + 
  labs(title = "Univariate scatter plot showing Absolute Error for each Test method", x = "Test") + 
  theme_minimal() + coord_flip()

peplot
```


############################################################################################################################
#Part C
#############################################################################################################################

# Filter the data
```{r}
PerceptionExp56to73 <- PerceptionExp %>% filter(Subject >= 56 & Subject <= 73) %>% mutate(Display = as.character(Display) %>% as.factor())
```


# qa
```{r}
Perception.agg <- PerceptionExp56to73 %>%
  dplyr::group_by(Test, estimateIndicator, Display) %>% # group the data by Test and EstimatedIndicator
  summarise(countoferror = n(), .groups = 'drop') %>% # find the count of error
  mutate(EstimateIndicator = as.factor(estimateIndicator))

barplot <- ggplot(Perception.agg, aes(x = reorder(factor(Test), countoferror), y = countoferror, fill = EstimateIndicator )) +
  geom_col(position = "dodge", colour = "black") +  # use the position parameter to create the grouped bar charts
  
  # This is for color blind people
  scale_fill_manual(values = c("#a6cee3", "#1f78b4","#b2df8a")) #+ scale_fill_brewer(palette = "Pastel1")


barplot <- barplot + theme(axis.text.x = element_text(face="bold.italic", 
                                   size=10, 
                                   angle=90),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=10
                                    ),  
        panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + labs(title = "A bar graph showing the comparison of equal, overestimated and underestimated responses over the different Test for Display 1 and 2", 
                 x = "Test")

barplot <- barplot + facet_wrap(~ Display)
barplot
```

########################################################################################################################################
# 1D
########################################################################################################################################

# Filter the data by Display and Test 
# Also derive a column anomalousresponse which identifies outlier observation in our data set
```{r}
display1verticaldistancenonaligned <- PerceptionExp %>% 
  filter(Display == "1" & Test == "Vertical Distance, Non-Aligned") %>% 
  mutate(AnomalousResponse = case_when(Response < 1 ~ "False", 
                                       Response == 1 ~ "True"
                                       ) %>% as.factor() # after transforming, change it to a factor
         )
```

# beeswarm plot
```{r}
library(ggbeeswarm)

ggswarmPlot <- ggplot(display1verticaldistancenonaligned, aes(x = Response, y = Subject, fill = AnomalousResponse)) + 
  geom_beeswarm(pch=21) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=10),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=10
                                    ),  
        panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + labs(title = "A beeswarm plot visualizing the subset of Outliers for Vertical Distance, Non-aligned for Display 1") + scale_fill_manual(values = c("#1f78b4", "#b2df8a")) 
 
ggswarmPlot
```


###############################################################################################################################################
# Question 2. Astronomical data set
#######################################################################################################################################

# Load the data set
```{r}
MessierData <- read.csv(file = "../HomeWork3/datasets/MessierData.csv", sep = "," , header = TRUE )
```

# determine any missing values
```{r}
sum(is.na(MessierData))
```

# Remove the missing missing values
```{r}
MessierData <- na.omit(MessierData)
```


# Rename all columns
```{r}
MessierData <- MessierData %>%
  rename(Messier = Messier.., 
         NGC = NGC.., 
         Kind = Kind, 
         Type = Type, 
         Distance.LY = Distance..LY., 
         Size = Size....., 
         RA.hm = R.A....h.m.., 
         Dec = Dec.........)
```

# Let's normalize the Distance, Size, Apparent.Magnitude variables
```{r}
library(scales)
MessierData <- MessierData %>% 
  mutate(Size.n = rescale(Size, to=c(0,1)), 
         Distance.LY.n = rescale(Distance.LY, to=c(0,1)), 
         Magnitude.n = rescale(Apparent.Magnitude, to=c(0, 1))
         )
```

# qa
```{r}
ggplot(MessierData, aes(x = Messier, y = Distance.LY + Size + Apparent.Magnitude)) + 
  geom_line() + 
  facet_wrap(~Distance.LY + Size + Apparent.Magnitude, scales = "free_y", ncol = 1)
```

####################################################################################################################################
#How would you approach 2(a)
# A scatter plot of one or more properties of the objects aganist the Messeir Number
################################################################################################################################

# Graph the Scatter plot of the log 10 of Distance aganist Messier Number
```{r}
splot <- ggplot(MessierData, aes(x = Messier, y = log(Distance.LY))) + 
  geom_point(size = 1.5) + 
  stat_smooth(method = lm, se = FALSE, colour = "black") + # se for disabling the CI, colour to change the regression line color 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + labs(title = "A Scatter plot of the Log 10 of Distance aganist the Messier Number", x = "Messier Number", y = "Log 10 of Distance.LY") 


splot
```

# Graph the Scatter plot of Size aganist Messier Number
```{r}
splot <- ggplot(MessierData, aes(x = Messier, y = Size)) + 
  geom_point(size = 1.5) + 
  stat_smooth(method = lm, se = FALSE, colour = "black") + # se for disabling the CI, colour to change the regression line color 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
    labs(title = "A Scatter plot of Size aganist the Messier Number", x = "Messier Number", y = "Size") 

splot
```

# Graph the Scatter plot of Apparent.Magnitude aganist Messier Number
```{r}
splot <- ggplot(MessierData, aes(x = Messier, y = Apparent.Magnitude)) + 
  geom_point(size = 1.5) + 
  stat_smooth(method = lm, se = FALSE, colour = "black") + # se for disabling the CI, colour to change the regression line color 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
    labs(title = "A Scatter plot of Apparent Magnitude aganist the Messier Number", x = "Messier Number", y = "Apparent Magnitude") 

splot
```





################################################################################################################################################## 2(b)
#################################################################################################################################



# beeswarm plot
```{r}
library(ggbeeswarm)

ggswarmPlot <- ggplot(MessierData, aes(x = fct_reorder(factor(Kind), log(Distance.LY)), y = log(Distance.LY))) + 
  geom_beeswarm(size = 1.5, alpha = 0.8) + 
  #expand_limits( y =0) +
  # theme_minimal() + # remove the theme
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + labs(title = "A beeswarm of Kind vs Log10(Distance)", x = "Kind", y = "log10 of Distance.LY") 
 
ggswarmPlot
```


#############################################################################################################################################
#2(C) Scatter plot
#############################################################################################################################################


# Graph the Scatter plot of the original variables
```{r}
splot <- ggplot(MessierData, aes(x = log(Distance.LY), y = Apparent.Magnitude)) + 
  geom_point(size = 1.5) + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + labs(title = "A Scatter plot of Log 10 of Distance aganist Apparent Magnitude", x = "Log 10 of Distance", y = "Apparent Magnitude") + scale_x_continuous(breaks = c(  7.5, 10, 12.5, 15))

splot
```

####################################################################################################
# 2(d)
###################################################################################################


# Graph the Scatter plot of the original variables
```{r}
splot <- ggplot(MessierData, aes(x = log(Distance.LY), y = Apparent.Magnitude)) + 
  geom_point(size = MessierData$Size, color = "#636363", fill = "lightskyblue", alpha = 0.4, shape = 21, stroke = 1.5) + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + labs(title = "A Scatter plot of Log 10 of Distance aganist Apparent Magnitude based on the Angular Size", x = "Log 10 of Distance", y = "Apparent Magnitude") + scale_x_continuous(breaks = c(  7.5, 10, 12.5, 15))

splot
```

############################################################################################################################################
#3 Montana Population data set
############################################################################################################

# load the data set
```{r}
montanapopds <- read.csv(file = "../HomeWork3/datasets/MontanaPopulationData.csv", sep = "," , header = TRUE )
```

# determine the number of observations
```{r}
dim(montanapopds)
```
# top 6 rows of the data set
```{r}
head(montanapopds)
```

# determine any missing values
```{r}
sum(is.na(montanapopds))
```

# structure of the data set
```{r}
str(montanapopds)
```

###########################################################################################################
#3 (a) Line graph
########################################################################################################
```{r}
lineplot <- ggplot(montanapopds, aes(x = Year, y = log(Population))) +
  geom_line() + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + labs(title = "Line graph of Log 10 of Population Vs Year", x = "Year", y = "Log 10(Population)") +
  scale_x_continuous(breaks = c(1890,  1910,  1930, 1950,  1970,  1990,  2010))
lineplot
```
#######################################################################################################################################
# 3 (b)
#######################################################################################################################################

# transform the dataset and add the previous year population variable
# and also derive the Percentage change of the population
```{r}
montanapopds <- montanapopds %>%
  mutate(prevYearPopulation = lag(Population) - 1) %>%
  mutate(perct.change.of.pop = round(((Population - prevYearPopulation)/prevYearPopulation) * 100 , 2))
```

# Line graph of Percentage Change of the Population Vs Year
```{r}
lineplot <- ggplot(montanapopds, aes(x = Year, y = perct.change.of.pop )) +
  geom_line(na.rm = TRUE) + # exclude an missing rows by using na.rm = TRUE
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + labs(title = "Line graph of Percentage Change of the Population Vs Year", x = "Year", y = "% Change in Population") +
  scale_x_continuous(breaks = c(1900,  1920, 1940,  1960,  1980,  2000)) + theme_minimal()
lineplot
```
##############################################################################################################################################
# 4 Air quality data set
##############################################################################################################################################

# Load the Air Quality data set
```{r}
airqualityds <- read.csv(file = "../HomeWork3/datasets/AirQuality.csv", sep = "," , header = TRUE )
```
# determine any missing values
```{r}
sum(is.na(airqualityds))
```

```{r}
airqualityds <- na.omit(airqualityds)
```


# structure of the air quality ds
```{r}
str(airqualityds)
```

# graph the scatter plot
```{r}
aqplot <- ggplot(airqualityds, aes(x = Wind, y = Solar.R)) + 
  geom_point(size = 1.3) + 
  stat_smooth(method = lm, se = FALSE, colour = "black") + # se for disabling the CI, colour to change the regression line color 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
    labs(title = "A Scatter plot of Wind Vs Solar.R") 

aqplot
```
###################################################################################################################################
# 4(b) Boxplot that will show the distribution of Wind and Solar.R
####################################################################################################################################

# pivot the datavalues
```{r}
airqualitycds <- airqualityds  %>%
  pivot_longer(c("Ozone", "Solar.R", "Wind", "Temp"), names_to = "Measurement", values_to = "Values") %>%
  mutate(Measurement = as.factor(Measurement))
```

# filter airquality data set by the Measurement values of Solar.R and Wind
```{r}
airquality.Wind.Solar.R <- airqualitycds %>%  filter(Measurement == "Solar.R" | Measurement == "Wind")
```

# graph the box plot
```{r}
bplot <- ggplot(airquality.Wind.Solar.R, aes(x = Measurement, y = Values)) + 
  geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3 , fill = "white") +
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
    labs(title = "Box plots comparing the distributions of Solar.R and Wind") 

# bplot <- bplot + geom_jitter()
bplot
```
#########################################################################################################################################
# 4(c) Box plots comparing the distributions of Ozone, Solar.R, Temp and Wind
#########################################################################################################################################


# graph the box plot
```{r}
bplot <- ggplot(airqualitycds, aes(x = Measurement, y = Values)) + 
  geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3 , fill = "white") +
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
    labs(title = "Box plots comparing the distributions of Solar.R and Wind") 


bplot
```

#########################################################################################################################################
# 4(d) Comparisson of Wind and Solar using the QQ plot
##########################################################################################################################################

# Sort the data set, airqualityds, 
```{r}
library(magrittr) # To access the dollar pipe, you need this library
library(scales)
sortedWindSolar <- airqualityds %$% # The dollar sign gives you access to the individual columns not 
  data.frame(wind = sort(Wind), 
             solar = sort(Solar.R)
             )
```

# normalize the Wind and Solar variables
```{r}
sortedWindSolar <- sortedWindSolar %>%
  mutate(windt = rescale(wind, to=c(0,1)
                         ), 
         solart = rescale(solar, to=c(0,1)
                          )
         )
```


# graph the qq plot
```{r}
xplot <- ggplot(sortedWindSolar, aes(x = windt, y = solart)) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0) +
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
    labs(title = "Q-Q comparing the distributions of Solar Vs Wind", x = "wind", y = "solar")
xplot
```

#############################################################################################################################################
# Question 5 Extra Credit
############################################################################################################################################

# Load the data set
```{r}
generalskillsurveyds <- read.csv(file = "../HomeWork3/datasets/General Skills Survey - Summary Report.csv", sep = "," , header = TRUE )
```

# get information about the data set
```{r}
str(generalskillsurveyds)
```

# clean up the variable names
```{r}
skills.surveyds <- generalskillsurveyds %>%
  rename(section = Section.., 
         question = Q.., 
         QnType = Q.Type, 
         QnTitle = Q.Title, 
         Bonus = Bonus., 
         Responses = X..Responses
         )
```

# get information about the data set, skills.surveyds
```{r}
str(skills.surveyds)
```

# drop section, Bonus, Difficulty variables from the data since they have missing values.
```{r}
skills.surveyds <- skills.surveyds %>% 
  select(-section, -Bonus, -Difficulty)
```

# determine if there is any missing values
```{r}
sum(is.na(skills.surveyds))
```

# students with no experience
```{r}
skills.noexp <- skills.surveyds %>% filter(Answer.Match == "N/A")
```

# Replace the null of values of QnTitle with string Visualization
```{r}
skills.surveyds <- skills.surveyds %>% mutate(QnTitle = replace(QnTitle, QnTitle == '', 'Visualization'))
```


# QnTitle with highest number of responses
```{r}
quest.num.responses <- skills.surveyds %>% 
  group_by(QnTitle) %>% 
  summarize(numresponses = sum(Responses)) %>%
  mutate(numresponses = sort(numresponses , decreasing = TRUE)) %>% # Order the aggregated data set by numresponses
  mutate(QnTitle = as.factor(QnTitle))
```

# bar graph of the number of responses vs QnTitle
```{r}
bplot <- ggplot(quest.num.responses, aes(x=fct_reorder(QnTitle, 
                                                        numresponses, .desc = FALSE
), 
y = numresponses, fill = QnTitle
)
) +
  geom_col(colour = "black") + 
  expand_limits(y=0) + 
  scale_fill_manual(values = c("#a6cee3", "#1f78b4","#b2df8a","#fbb4ae", "#d95f02"
                               )
                    ) + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
        ),  
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
        ) # format the title
  ) + 
  labs(title = "Number of Responses per Question Title", x = "QnTitle", y = "Number of Responses")  + 
  scale_fill_discrete(limits=c("Commercial tool Importance", "Commercial Software Tool Experience", "Data sets of interest", "Programming Experience", "Visualization")) + # controlling the order of the legend. 
  coord_flip() + theme_bw()

bplot

```

# graph the box plot
```{r}
bplot <- ggplot(skills.surveyds, aes(x = as.factor(QnTitle), y = Responses)) + 
  geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3 , fill = "white") + expand_limits(y = 0) +
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
    labs(title = "Box plots comparing the distributions of the Question Titles", x = "Question Title")  + coord_flip() + theme_bw()


bplot
```

# Univariate scatter plot
```{r}
bplot <- ggplot(skills.surveyds, aes(x = as.factor(QnTitle), y = Responses)) + 
  geom_point() + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=12),  
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=12
                                    ),  
        # panel.grid.major.x = element_blank(), # Hide the horizontal grid lines
        # panel.grid.minor.x = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
        ) + 
    labs(title = "Univariate Scatter plot showing the distribution of Responses aganist the different Question Titles", x = "Question Title")  + coord_flip() + theme_bw()


bplot
```
 # Bar graph visualizing the whose response was N/A
```{r}
skills.noexp.agg <- skills.noexp %>% 
  mutate(QnTitle = as.factor(QnTitle)) %>%
  group_by(QnTitle) %>% # group and find the sub totals
  summarize(TotalResponses = sum(Responses))
  
bplot <- ggplot(skills.noexp.agg, aes(x=fct_reorder(QnTitle,
                                                        TotalResponses, .desc = FALSE
                                                       ),
                                         y = TotalResponses, fill = QnTitle
                                         )
                ) +
  geom_col(colour = "black") +
  expand_limits(y=0) +
  scale_fill_manual(#breaks = legend_ord,
                    values = c("#a6cee3", "#1f78b4","#b2df8a"
                               )
                    ) +
  theme(axis.text.x = element_text(face="bold.italic",
                                   size=12),
        axis.text.y = element_text(face="bold.italic",

                                   size=12
        ),
        plot.title = element_text(hjust = 0.5,
                                  size = 14,
                                  face = "bold.italic"
        ) # format the title
  ) +
  labs(title = "Number of Responses per Question Title where students answered N/A", x = "QnTitle", y = "Number of Responses")  +
  coord_flip() + 
  theme_bw()

bplot
```
 