---
title: "DSC 465 HomeWork1 Assignment"
author: "Ronaldlee Ejalu"
date: "4/1/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


#load the necessary packages
```{r}
#library(plyr) # for data wrangling
library(dplyr)     # for data wrangling
library(lubridate) # for dealing with dates
library(ggplot2)   # for dealing with visualization
```

#set the working directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC 465/HomeWork/HomeWork1")
```

#Load the Intel Stock data set
```{r}
intelstock <- read.csv(file = "../HomeWork1/datasets/Intel-1998.csv", sep = "," , header = TRUE )
```

#view the first 6 records of the data set
```{r}
head(intelstock)
```

#Find any missing values
```{r}
sum(is.na(intelstock))
```

#derive weekday
```{r}
intelstockds <- intelstock %>% 
  select(date = Date, 
         trainingday = Trading.Day, 
         open = Open, 
         high = High, 
         low = Low, 
         close = Close, 
         volume = Volume, 
         adjclose = Adj.Close, date1 = Date,
         ) %>%
  transmute(date = mdy(date), 
            date1 = date1,
            trainingday = trainingday, 
            open=open, 
            high=high, 
            low=low, 
            close=close, 
            volume=volume, 
            adjclose=adjclose,
            range = (high - low)
            ) %>%
  mutate(wday = wday(date, label = TRUE, abbr = TRUE), date1 = as.Date(date)
                                          
         )
```


# An ordinary line graph of closing price vs date
```{r}
ggplot(data = intelstockds, aes(x = date, y = close)) + 
  geom_line(size = 1.5, alpha = .6, position = position_dodge(0.2)) + 
  ylim(0, max(intelstockds$close)) + # have the y range start from zero. 
  annotate("text", 
           x = mean(range(intelstockds$date)), 
           y = Inf, 
           label = "closing price vs date", 
           vjust = 1.5, 
           size = 4.5, 
           colour = "red"
           )
```


#Bar graph of Volume Vs the exact Date
```{r}
#specify breaks as a Date vector
datebreaks <- seq(as.Date("1998-01-01"), as.Date("1999-01-01"), by = "1 month")
p <- ggplot(intelstockds, aes(x = date, y = volume)) + 
  geom_col(width = 0.5, position = "dodge") + 
  ggtitle("Volume vs date")

p <- p + scale_x_date(breaks = datebreaks) + 
  scale_y_continuous(labels = scales::comma) + # format the y axis and include , in the volume values
  theme(axis.text.x = element_text(face="bold", 
                                          color="#993333"
                                          #,  size=14
                                          , angle=90)
               , axis.text.y = element_text(face="bold",
                                             color="#993333"
                                            #, size=14
                                            , angle=45)
               ) 

p  
```

# normalize the volume 
```{r}
library(scales)
intelstockds <- intelstockds %>%
  mutate(norm.volume = scale(volume))
```

# structure of intelstockds
```{r}
str(intelstockds)
```

# summary statistics of norm.volume
```{r}
summary(intelstockds$norm.volume)
```


# determine the bin width
# using the formula 2 * IQR/cube root of n
```{r}
# Q3 = 0.3017  
# Q1 = -0.5989
# IQR = Q3 - Q1
# bin width  = 2 * IQR/cube root of n
```

# number of observations 
```{r}
dim(intelstockds)
```
Filt


#Creating a histogram of the daily stock Volume
```{r}
ggplot(intelstockds, aes(x=volume)) + 
  geom_histogram(bins = 20) + 
  ggtitle("Histogram of daily stock volume") +
  scale_x_continuous(labels = scales::comma)
```



# Remove entries with the means greater than 150086000
```{r}
intelstockds.sortedf  <- intelstockds %>% filter(volume <= 151875200)
```


#Repeat or corrections for 1C
```{r}
ggplot(intelstockds.sortedf, aes(x=log(volume))) + 
  geom_histogram(bins = (log2(252)) + 1) + 
  expand_limits(y=0) + 
  theme(axis.text.x = element_text(face="bold.italic", 
                                   
                                   size=10
                                   ), 
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=10
                                    )
        ,
        plot.title = element_text(hjust = 0.5, 
                                  size = 14, 
                                  face = "bold.italic"
                                  ) # format the title
         
        )  + labs(x = "natural log of volume", title = "Histogram of daily stock volume")
```



# Scatter plot that graphs the Volume on the x-axis and
# the daily price on the y-axis. 
```{r}
ggplot(intelstockds, aes(x = volume, y = range)) + 
  geom_point(size = 1.5) + # using the size aesthetic to control the size of the points.
  ggtitle("Scatter plot of daily price range vs Volume ") + 
  scale_x_continuous(labels = scales::comma) 
```

###########################################################################################################################################
#Question 2
###########################################################################################################################################

#Load the data set
```{r}
perceptionExperiment <- read.csv(file = "../HomeWork1/datasets/PerceptionExperiment1.csv", sep = "," , header = TRUE )
```

#View the first 6 rows
```{r}
head(perceptionExperiment)
```

#derive a new column that contains the amount of error. 
```{r}
perceptionExperiment <- perceptionExperiment %>%
  mutate(error = Response - TrueValue, 
         absoluteError = abs(Response - TrueValue), Test = as.factor(Test)
         )
```

#a histogram of the overall distribution of Error
```{r}
ggplot(perceptionExperiment, aes(x=error)) + 
  geom_histogram(bins = 20) + 
  ggtitle("Histogram of the overall distribution error") #+
  #scale_x_continuous(labels = scales::comma)
```

# To be discussed in class. 
# a bar graph of the median Error by Test
```{r}
library(tidyverse)
ggplot(perceptionExperiment, aes(x = fct_reorder(Test, error), y = error)) + 
  geom_col() + 
  theme(axis.text.x = element_text(face="bold", 
                                          color="#993333"
                                          ,  size=10
                                          , angle=90
                                   ), 
        axis.text.y = element_text(face="bold", 
                                   color="#993333"
                                   )
  ) + 
  labs(title = "A bar graph of median error by Test", x = "Test") + 
  coord_flip()
```

#Repeat of 2b
# a bar graph of the median Error by Test

```{r}
library(tidyverse)
p <- ggplot(perceptionExperiment, aes(x=fct_reorder(Test, desc(error)), y=error, fill=Test))
p <- p + stat_summary(fun.y="median", geom="bar") + theme_grey(base_size = 14)
p <- p +  theme(axis.text.x = element_text(face="bold" 
                                           ,  size=10
                                           , angle=90
                                   ),
        axis.text.y = element_text(face="bold"
                                   ),
        panel.grid.major.x = element_blank(), # Hide the vertical grid lines 
        panel.grid.minor.x = element_blank(),
        
        panel.grid.major.y = element_blank(), # Hide the horizontal grid lines
        panel.grid.minor.y = element_blank(), 
        
        # Change the background color (fill) and border (colour)
        # and border thickness (size)of the legend box
        # We control the title text color, font size and font face
        
        legend.background = element_rect(fill = "white", colour = "grey", size = 1), 
        legend.title = element_text(colour = "blue", face = "bold", size = 14),
        legend.text = element_text(colour = "black"), 
        legend.key = element_rect(colour = "black", size = 0.1)
  ) + 
  labs(title = "A bar graph of median error by Test", x = "Test")
p <- p + scale_fill_discrete(limits = c("Slope", "Volume", "Area",  # controlling the order of the legend
                                   "Veritcal Distance, Aligned", 
                                   "Vertical Distance, Non-Aligned", 
                                   "Angle", "Length, Non-Aligned", "Color Value")) 
p + scale_y_continuous(breaks = c(0.1, 0, -0.1)) + theme_bw()  + coord_flip()

```

#get variables of interest to derive the standard deviation
```{r}
perceptionExperimentsdErrorbyTestdf <- perceptionExperiment %>%
  select(Test = Test, 
         #TestNumber = TestNumber,  
         error = error
         )
```

#Function to calculate the mean and the standard deviation for each group
```{r}
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}
```

#Summarize the data set
```{r}
perceptiondf <- data_summary(perceptionExperimentsdErrorbyTestdf, 
                             varname = "error", 
                             groupnames = c("Test")
                             )
perceptiondf$Test <- as.factor(perceptiondf$Test)
```

#plot a bar graph of Standard deviation of the Error by Test
```{r}
p <- ggplot(perceptiondf, aes(x = fct_reorder(Test, desc(sd)), y = sd)) + 
  geom_col(colour="black") +  
  expand_limits(y=0) + # Make the y-axis go to zero. 
  theme_grey(base_size = 14) +
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=10, 
                                   angle=90),
        
        axis.text.y = element_text(face="bold.italic", 
                                   #color="#993333"
                                   ), 
        panel.grid.major.x = element_blank(), # Hide the vertical grid lines 
        panel.grid.minor.x = element_blank(),
        
        panel.grid.major.y = element_blank(), # Hide the horizontal grid lines
        panel.grid.minor.y = element_blank(), 
        legend.position = "none"
        ) + 
  labs(y  = "Standard deviation of Error", x = "Test", title = "Standard deviation of Error\nby Test")    
p <- p +  coord_flip() + scale_y_continuous(breaks = c(0.00, 0.05, 0.10, 0.15, 2.0))
p
```

# a bar graph of the Absolute Error by Test
```{r}
library(tidyverse) 
perceptionExperiment <- perceptionExperiment %>% mutate(Test = as.factor(Test)) %>% mutate(results = fct_reorder(Test, absoluteError)) # ordering the data

p <- ggplot(perceptionExperiment, aes(x = fct_reorder(Test, absoluteError), y = absoluteError)) + 
  geom_col() + 
  expand_limits(y=0) + # Make the y-axis go to zero. 
  theme_grey(base_size = 14) +
  theme(axis.text.x = element_text(face="bold.italic", 
                                   size=10, 
                                   angle=90),
        
        axis.text.y = element_text(face="bold.italic", 
                                   ), 
        panel.grid.major.x = element_blank(), # Hide the vertical grid lines 
        panel.grid.minor.x = element_blank(),
        
        panel.grid.major.y = element_blank(), # Hide the horizontal grid lines
        panel.grid.minor.y = element_blank(), 
        legend.position = "none"
        ) 
p <- p + 
  labs(x = "Test", title = "Absolute Error Vs Test", y = "Absolute Error") +
  coord_flip() + scale_y_continuous(breaks = c(0, 25, 50, 75, 100))
p
```

#####################################################################################################################################################################
#Question number 3
#####################################################################################################################################################################

#Load the Infant data set.
```{r}
infantdata <- read.csv(file = "../HomeWork1/datasets/InfantData.csv", sep = "," , header = TRUE )
```

#find an missing variables
```{r}
sum(is.na(infantdata))
```

#Convert sex to a factor
```{r}
infantdatads <- infantdata %>%
  transmute(Sex = Sex, 
            HeightIn = HeightIn, 
            WeightLbs = WeightLbs
  ) %>%
  mutate(Sex = as.factor(Sex), 
         color = as.factor(case_when(Sex == "M" ~ "#a6cee3", 
                                     Sex == "F"  ~  "#b2df8a")
                           )
  )
```


# Scatter plot of Weightlbs Vs HeightIn
```{r}
colr  <- as.character("blue", "green")
p <- ggplot(infantdatads, 
       aes(x = HeightIn, 
           y = WeightLbs
           , shape = Sex
           )
       ) + 
  geom_point(color=infantdatads$color) + # set the size of the points to 1.5
  
  theme_grey(base_size = 14)# +  scale_fill_manual(values = c("#a6cee3", "#1f78b4"))
p <- p + theme(axis.text.x = element_text(face="bold.italic", 
                                   
                                   size=10
                                   ), 
        axis.text.y = element_text(face="bold.italic", 
                                   #color="#993333", 
                                   size=10
                                   ),
        panel.grid.major.x = element_blank(), # Hide the vertical grid lines
        panel.grid.minor.x = element_blank(),

        panel.grid.major.y = element_blank(), # Hide the horizontal grid lines
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0.5,
                                    size = 14,
                                    face = "bold.italic"
          )
        )  + labs(title = "Scatter plot of Weight in Lbs Vs Height in inches", x = "Height in inches", y = "Weight in Lbs") + scale_color_manual(values = levels(infantdatads$color))
p <- p + scale_x_continuous(breaks = seq(15.0, 25.0, by = 2.5)) + theme_bw()
p
```


# Scatter plot of Weightlbs Vs HeightIn
# Adding separate single trend lines
```{r}
p <- ggplot(infantdatads, 
       aes(x = HeightIn, 
           y = WeightLbs
           , colour = Sex
           , shape = Sex
           )
       ) + 
  geom_point() + # set the size of the points to 1.5
  stat_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter plot of Weight in lbs Vs Height in inches\nwith separate trend lines", x = "Height in inches", y = "Weight in Lbs") + 
  theme_grey(base_size = 14) 
p <- p + theme(axis.text.x = element_text(face="bold.italic", 
                                   
                                   size=10
                                   ), 
        axis.text.y = element_text(face="bold.italic", 
                                   
                                   size=10
                                   ),
        panel.grid.major.x = element_blank(), # Hide the vertical grid lines
        panel.grid.minor.x = element_blank(),

        panel.grid.major.y = element_blank(), # Hide the horizontal grid lines
        panel.grid.minor.y = element_blank(),
        )  + theme_bw()
p
```
